### 2011-03-09 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/evaluation.tex
New version for IJCAI workshop

We have implemented our switching continual approach in the MAPSIM
environment~\cite{brenner:nebel:jaamas09}. Our implementation is able
to use several underlying planning systems. We have also extended the
MAPSIM system so that it can parse DTPDDL, and perform successive
estimation of the underlying belief-state using
\system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
we use our own version of Fast Downward~\cite{fast-downward} for
sequential sessions. We have extended that system to support actions
with success probabilities. In our evaluation Fast Downward is run
with the cyclic causal graph heuristic using an A* search or weighted
A* (with weight 5). We use A* for easy problems and WA* for more
difficult problems where A* is ineffective. Contingent sessions use
our own forward search procedure.

\Omit{ We then perform multiple tests with different limits for the
  belief-space size in contingent sessions.  Higher limits should
  cause longer planning times but be beneficial to plan quality as
  more contingencies can be taken in to account by the POMDP planner.
}

In order to evaluate our system, we have also implemented a baseline
approach in MAPSIM. Rather than invoking a contingent session when a
switching action $\switchAction$ is scheduled for execution, the
baseline simply executes one action that can trigger a sensing outcome
determined by the precondition of $\switchAction$, and replans in the
resulting belief-state.

%% executes switching actions and replans in the
%% resulting belief state.

%% but instead of
%% creating an observation problem for the decision theoretic planner it
%% will just execute one sensing action -- assuming that this action will
%% confirm its assumption.

To test our approach we use a robot exploration domain based on a
scenario from our physical robotic system. Here, a robot is exploring
an office or living environment, and trying to report the locations of
objects to their owner. Basic types are {\tt rooms}, {\tt places} and
{\tt objects}. Places are topological map nodes that occur in rooms.
Objects, and also the robot, are always located at a place. The robot
can move around the rooms via connections between places given by a
{\tt connected} predicate. Each room has a, possibly unknown, {\em
  category} (e.g. kitchen, office, living room). Certain objects are
more likely to be located in rooms of a particular category.  For
example, a box of cornflakes is more likely to be located in the
kitchen than the office. The robot can look for an object at a place
by executing a {\tt look-for-object} action. Executing {\tt
  look-for-object} results in a noisy sensing outcome, sometimes
indicating a positive perception if the object is there. We have that
some objects are harder to detect than others. Finally, if the robot
is in the presence of a human, it can ask what type of room it is
currently in, however conducting a dialogue is more costly than
running a vision algorithm (cost of 8 vs costs of 3).

%%  Therefore, an instantaneous positive detection of an object is
%% not proof positive of it being there.

We evaluated our switching planner against the baseline in several
tasks with the number of rooms ranging from 3 to 6. We examine the
impact of sensing reliability, experimenting using sensor models that
are: ({\em reliable}) $0.9$ chance the object is perceived if it is
present, ({\em semi-reliable}) chance is 0.65, and ({\em noisy})
chance is 0.4. The continual planning times are reported in
Figure~\ref{fig:results-time}, and the quality data in
Figure~\ref{fig:results-quality}. Here, in switching runs we examine
the performance of our system where $b_0$ allocates non-zero
probability to between 20 and 200 abstract states during contingent
sessions. For Item~$a$-$e$ the objective is to find one or more
objects and report their position to the user. In some of the tasks we
configured the environment so that no plan exists, by excluding the
desired object from the environment. Item~$f$ reports for tasks
requiring indirect sensing, where the robot must relocate to a
particular room category. For Item~$f$ there is no human in the
environment, and therefore the baseline cannot be used, because there
are no sensing schemata that directly sense room categories. In our
evaluation we run 50 simulations in each configuration, and have a
timeout on each simulation of 30 minutes (1800 seconds). In our domain,
rewards encapsulate a goal condition, and therefore we find it
intuitive to report average costs of the plans and the success rate in
problems that admit a solution (i.e., positive reward scaled by a
constant factor).

%% Because the switching utility model

%% Using a satisficing, optimistic serial planner in sequential sessions
%% makes optimising for the expected reward difficult. Only reasoning
%% about single traces, it is optimistic wrt the remaining costs to a
%% goal and pessimistic regarding the probability of reaching it. By
%% setting the rewards to non-extremal values has little effect on that
%% system. Therefore, we do not consider expected utility in our
%% experiments. Instead, we report average costs of the plans and the
%% success rate (as a ratio between solved and solvable problems).

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora2-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora3-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora4-time}\hfill
  \vspace{2mm}
  \includegraphics{dora56-time}\hfill
  \vspace{2mm}
  \includegraphics{dora-cat-time}\hfill
  \caption{Average runtime}
  \label{fig:results-time}
\end{figure}

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora2-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora3-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora4-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora56-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora-cat-quality}\hfill
  \caption{Average plan costs and number of successful runs.}
  \label{fig:results-quality}
\end{figure}

%% Figure \ref{fig:results-quality} shows the average costs of the
%% executed plans as well as the percentage of solvable tasks that were
%% actually solved by the planner.

Examining success ratios and plan costs, where sensing is reliable
there is little to be gained by using the contingent planner, as the
greedy sensing approach of the baseline is sufficient. Not
surprisingly, as sensing degrades contingent planning pays off.  Also,
we find that time spent in contingent planning increases steeply as
the abstraction $\bstate_0$ becomes more refined.  That refinement
seems to be paying off in terms of the success ratio, particularly for
tasks $d$ and $e$, where we had sequential sessions using weighted
$A^*$ (rather than $A^*$). For less refined initial configurations,
the increase cost of contingent planning is compensated by a decrease
in Fast Downward planning times. The relatively high success rate
irrespective of the level of refinement in the initial configuration
indicates the effectiveness of using conditional entropy to guide
abstraction refinement in our setting.

%% :
%%while the resulting plans are still
%% longer on average, the impact on the number of solved tasks was much
%% smaller than for the baseline system.

%% Less aggressive abstraction of the initial configuration results in
%% longer runtimes, with little to be gained in terms of plan quality. We
%% conclude that in our scenario it is worthwhile being assumptively
%% aggressive.

%% We find that abstraction has little impact on plan
%% costs and success rate in our scenario. Increasing the size of the
%% initial abstract belief beyond 50 states rarely pays off, because
%% while additional information may facilitate better plans, the indirect
%% sensing tests report the same result. Here, we only really see an
%% impact in terms of the runtime of contingent planning.

% We believe that a part of the improvement is due to the segmentation of
% the plan into several subtask, essentially performing hierarchical
% planning. Especially when the continual planner performs badly this is
% a huge gain.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "moritz_2011"
%%% End:

>>>>

We have implemented our switching approach in the MAPSIM
environment~\cite{brenner:nebel:jaamas09},
using \system{dlib-ml}~\cite{king:2009} for belief
revision. Sequential sessions use a modified version of Fast
Downward~\cite{fast-downward}, and DT sessions use our own contingent
procedure. Since most of the problems we consider are much larger than
any available DT planner can solve directly, for comparison purposes
we also implemented a simple dual-mode replanning baseline
approach. Here, when a switching action is scheduled for execution the
DT session plans to a single entropy reduction action, whose execution
can provide evidence regarding the truth value of a relevant
assumption.
%%
%%  -- i.e,. we say a fact can
%% be directly sensed, if there is a DTPDDL {\em sense} declaration that
%% mentions the fact in its precondition. 
%%
Control is then immediately
returned to a new sequential session.

We evaluate our approaches in robot exploration tasks from home and
office environments. Spatially, these consist of {\em rooms}
(office/kitchen/etc), and an underlying topological map over
smaller areas of space, called {\em places}, and connectivity between
those. The mobile {\em robot} and {\em visual objects} inhabit the
topological places. Objects indicate the category of space they
inhabit---e.g., spoons are likely to be in kitchens. By examining
view cones at places for particular objects, the robot is able to: (1)
categorise space at high (room) and low (place) levels, and (2) find
objects for the user, exploiting information about object
co-occurrence and room categories for efficiency. Also, in the
presence of a person, the robot can ask about the category of the
current room.

We compare switching to the baseline in several realistic
tasks, with the number of rooms ranging from 3 (12-places, 16-objects,
$|$states$|$ $>10^{21}$) to 6 (26-places, 21-objects, $|$states$|$
$>10^{36}$). We also compare those systems with near optimal policies
computed using Smith's {\sc zmdp} for small 2 room problems (4-places,
3-objects, $|$states$|$ $\simeq 5000$). Our evaluation considers 3
levels of reliability in sensing: {\em reliable} sensors have a $.1$
probability of a false negative, {\em semi-reliable} have a chance of
$0.3$ of false negative and $0.1$ of false positive, and {\em noisy}
sensors with probabilities of $0.5$ and $0.2$ respectively. Each
object class is assigned one sensor model---e.g. cornflakes may be
harder to detect than refrigerators. We performed several experiments
with different levels of reliability for sensing the target object(s),
while keeping sensing models for non-target objects constant.

\Omit{
All non-target objects remain unchanged so the planner can
exploit object co-occurrences using reliably sensed objects to
compensate for the increased sensor noise.
}

Our evaluation examines DT sessions with initial belief-states
admitting between 20 and 100 abstract states with non-zero
probability. We run 50 simulations in each configuration, and have a
timeout on each simulation of 30 minutes (1800 seconds)\footnote{All
experiments were conducted on a 2.66GHz Intel Xeon X5355 using one CPU
core.}. The continual planning times are reported in
Figure~\ref{fig:results-time}, and the quality data in
Figure~\ref{fig:results-quality}.
%%
For each task, the goal is to find one or more objects and report
their position to a user. Usually there is a non-zero probability that
no plan exists, as the desired object might not be present in the
environment.
% MOG: I didn't do  evaluations for those, so let's remove them
% Item~$f$ reports for tasks requiring indirect sensing,
% where the robot must relocate to a room with a particular category
% (e.g. kitchen). the {\em baseline} is not able to complete Item~$f$,
% so is omitted from that reporting.  
In these experiments we only allocate reward on the achievement of all
goals, therefore we find it intuitive to report average plan costs and
the success rates in problems that admit a complete solution (i.e.,
positive reward scaled by a constant factor). The exception occurs for
items f and g of Figure~\ref{fig:results-quality}, where we report
expected discounted rewards (not plan costs).

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora2-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora3-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora4-time}\hfill
  \vspace{2mm}
  \includegraphics{dora56-time}\hfill
  % \vspace{2mm}
  % \includegraphics{dora-cat-time}\hfill
  \caption{Average runtime}
  \label{fig:results-time}
\end{figure}

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-quality}\hfill
  % \vspace{2mm}
  % \includegraphics{pomdp-solvable-quality}\hfill
  \includegraphics{dora2-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora3-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora4-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora56-quality}\hfill
  \vspace{2mm}
  \includegraphics{pomdp-quality}\hfill
  % \vspace{2mm}
  % \includegraphics{dora-cat-quality}\hfill
  \caption{Average plan costs and number of successful runs.}
  \label{fig:results-quality}
\end{figure}

We find that if sensing is reliable, then little is gained using DT
sessions, as the greedy approach of the baseline is
sufficient. As sensing degrades DT sessions prove more useful. Here,
time spent on DT planning increases steeply as the abstraction becomes
more refined, which is compensated for by fewer planning sessions
overall. More detailed abstractions lead to a better overall
success rate, particularly for tasks $d$ and $e$.
%% Moreover,
%% for less refined abstractions, the increased cost of DT planning is
%% partially compensated for by a decrease in Fast Downward planning
%% times. 
Speaking to the effectiveness of our entropy heuristic for abstraction
refinement, we see relatively high success rates irrespective of
the level of refinement. Comparing finally to the best {\sc
zmdp} policy, although producing relatively costly plans, the
continual planners performed quite well, especially in terms success
rate. A key source of inefficiency here, is due to sequential sessions
always being optimistic, and refusing to abandon the search.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "aaai11"
%%% End:

### 2011-03-09 code/subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/relatex.tex
New version for IJCAI workshop

There has also been much recent work on scaling offline approximate
POMDP solution procedures to medium-sized instances. Recent
contributions propose more efficient belief-point sampling
schemes~\cite{kurniawati:etal:2010,shani:etal:08}, and factored
representations with procedures that can efficiently exploit
structures in those
representations~\cite{brunskill:russell:2010,shani:etal:2008}. Offline
domain independent systems scale to {\em logistics} problems with
$2^{22}$ states~\cite{shani:etal:2008}, taking over an hour to
converge, and around 10 seconds on average to perform each Bellman
backup. \citeauthor{brunskill:russell:2010} are able to solve problems
with approximately $10^{30}$ states, by further exploiting certain
problem features -- E.g., problems where no actions have negative
effects.
%%
Moving someway towards supporting real-time decision making, recent
online POMDP solution procedures have been developed which leverage
highly approximate value functions -- computed using an offline
procedure -- and heuristics in forward
search~\cite{ross:etal:2008}. These approaches are applicable in
relatively small problems, and can require
expensive \emph{problem-specific} offline processing in order to yield
good behaviours.
%%
A {\em very} recent and promising online approach for
larger POMDPs employs Monte-Carlo sampling to break the curse of
dimensionality in situations where goal reachability is {\em easily}
determined~\cite{silver:veness:2010}. 
%%
Our approach can also be thought of as an online POMDP solver that
uses a sequential plan to guide the search, rather than (e.g.,
Monte-Carlo) sampling. Also, compared to most online POMDP procedures, which
replan at each step, our approach involves relatively little replanning.

>>>>

There has also been much recent work on scaling offline approximate
POMDP solution procedures to medium-sized instances. Recent
contributions propose more efficient belief-point sampling
schemes~\cite{kurniawati:etal:2010,shani:etal:08}, and factored
representations with procedures that can efficiently exploit
structures in those
representations~\cite{brunskill:russell:2010,shani:etal:2008}. Offline
domain independent systems scale to {\em logistics} problems with
$2^{22}$ states~\cite{shani:etal:2008}, taking over an hour to
converge, and around 10 seconds on average to perform each Bellman
backup. \citeauthor{brunskill:russell:2010} are able to solve problems
with approximately $10^{30}$ states, by further exploiting certain
problem features -- E.g., problems where no actions have negative
effects.
%%
Moving someway towards supporting real-time decision making, recent
online POMDP solution procedures have been developed which leverage
highly approximate value functions----computed using an offline
procedure----and heuristics in forward
search~\cite{ross:etal:2008}. These approaches are applicable in
relatively small problems, and can require
expensive \emph{problem-specific} offline processing in order to yield
good behaviours.
%%
A {\em very} recent and promising online approach for
larger POMDPs employs Monte-Carlo sampling to break the curse of
dimensionality in situations where goal reachability is {\em easily}
determined~\cite{silver:veness:2010}. 
%%
Our approach can also be thought of as an online POMDP solver that
uses a sequential plan to guide the search, rather than (e.g.,
Monte-Carlo) sampling. Also, compared to most online POMDP procedures, which
replan at each step, our approach involves relatively little replanning.
